---
title: "Cow SNA"
author: "Rafael Pilliard-Hellwig"
date: "05/07/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Intro

The purpose of this notebook is to build a proof-of-concept for a social network analysis of cows. 

Let's start by loading our packages.

```{r pkgs, warning=FALSE, message=FALSE}
# load packages
library(tidyverse)
library(tidygraph)
library(visNetwork)
```

Next, let's load in the given data. The names are bit long, so although not generally advisable, let's attach the object to the namespace. This will mean we simply need to access the `HOBO` and `Intensec` objects directly.

```{r}
# Load the data. Takes about 10 seconds
system.time({
  load(here::here("data/full_10_month_analysis_result_summary_only_dashboard.Rda"))
  attach(dashboard_full_analysis)
  rm(dashboard_full_analysis)
})

# print the objects in the namespace
ls()
```

The HOBO data consists of trackers on cows that track their movement. The data has been pre-processed already, so let's start simple and simply explore one of the adjacency matrices for a given date. We'll consider this an undirected network, which will simply things greatly.

```{r}
# an adjacency matrix of cows that hang out together
class(HOBO$`paired lying total time`$`2020-07-13`)

# size
dim(HOBO$`paired lying total time`$`2020-07-13`)
```

Let's make this into a tidy graph:



```{r}
# build a tidygraph
(g <- as_tbl_graph(HOBO$`paired lying total time`$`2020-07-13`, add.colnames = TRUE, add.rownames = TRUE, directed = FALSE) )

# add measures of centrality and clusters
g <- g %>%
  activate(nodes) %>%
  mutate(eigen = centrality_alpha(weights = weight),
         authority = centrality_authority(weights = weight),
         power = centrality_power(),
         grp1 = group_fast_greedy(weights = weight),
         grp2 = group_infomap(weights = weight),
         grp3 = group_leading_eigen(weights = weight),
         grp4 = group_louvain(weights = weight),
         grp5 = group_walktrap(weights = weight, steps = 4))
```

This tidygraph can be visualized:

```{r}
# this workaround is needed to get the cow icon to work
txt <- RCurl::base64Encode(
  readBin(
    "www/cow1-black.png",
    "raw",
    file.info("www/cow1-black.png")[1, "size"]
  ),
  "txt"
)
```


```{r}
edges <- g %E>%
  as_tibble %>%
  arrange(desc(weight)) %>%
  group_by(from) %>%
  top_n(3) %>%
  mutate(width = scales::rescale(weight, to = c(1, 20)))
nodes <- g %N>% 
  as_tibble %>% 
  tibble::rowid_to_column("id") %>%
  mutate(label = name, 
         shape = "image",
         size = 10,
         group = grp5,
         #group = "cow",
         image = paste('data:image/png;base64', txt, sep = ','),
         size = scales::rescale(authority, to = c(15, 40)))
visNetwork(nodes, edges, width = "100%", height = "800px") 
```

Here, the colours represent the clusters, and the sizes represent the centrality of the cow. I've paired down the edges list a bit arbitrarilty just to make it manageable to visualize. A lot can be improved on this viz, but it's a start.


## Image Attribution

<div>Icons made by <a href="" title="Vitaly Gorbachev">Vitaly Gorbachev</a> from <a href="https://www.flaticon.com/" title="Flaticon">www.flaticon.com</a></div>














